---
title: "Machine Learning Project"
author: "Michael Sewerin"
date: "17 November 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

### Goal of the project
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.


```{r libraries}
library(ggplot2)
library(caret)
```

Download data:
```{r download}
fileUrlTraining <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
fileUrlTesting <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(fileUrlTraining, destfile = "pml-training.csv")
download.file(fileUrlTesting, destfile = "pml-testing.csv")
```

From looking at the csv-files we can see, that there are multiple possibilities for NA-Strings, so when importing the data we add them to na.strings

```{r import}
training <-  read.csv("pml-training.csv", na.strings = c("NA", "#DIV/0!", ""))
testing <- read.csv("pml-testing.csv", na.strings = c("NA", "#DIV/0!", ""))
```

Take a look at the data with head(training) - (too much to display here)
From this we can see that there are a lot of variables with missing (NA) data. We remove those that have only NA values
```{r na_remove}
training <- training[, colSums(is.na(training)) == 0]
testing <- testing[, colSums(is.na(testing)) == 0]
```

We also see that the first 7 variables are just IDs or names, so we can also remove them
```{r remove}
training <- training[, -c(1:7)]
testing <- testing[, -c(1:7)]
```

Now we can split training into a training and a test dataset.
```{r partition}
set.seed(1234)
trainingPartition <- createDataPartition(training$classe, p = 0.7, list = FALSE)
trainingSet <- training[trainingPartition,]
testingSet <- training[-trainingPartition,]
```

### Model 1: Classification Tree
First we try a Classification Tree model:
```{r model_class}
modClass <- train(classe ~ ., data = trainingSet, method = "rpart")
predClass <- predict(modClass, testingSet)
confusionMatrix(predClass, testingSet$classe)
```

The accuray is not very good (49%), so as a next step we try a random forest model

### Model 2: Random Forest

We calculate the model with 3 fold cross validation
```{r model_rf}
tc <- trainControl(method="cv",number=3)
modRF <- train(classe ~ ., data = trainingSet, methot = "rf", trControl = tc)
predRF <- predict(modRF, testingSet)
confusionMatrix(predRF, testingSet$classe)
sum(predRF != testingSet$classe)/length(testingSet$classe)
```

We see that this has an accuracy of 99.41% and an out of sample error rate of 0.59%. So we choose the random forest model as our final model.

###Predicting test data
Now we can use our final model to predict the classe for the imported test data:
```{r predict_test}
predRFTest <- predict(modRF, testing)
predRFTest
```